Disparities in Climate-Induced Health Outcomes in the Greater Toronto Area.

### 2022-07-18 outputs/paper/paper.qmd
Updates to paper

Exposure to NO (Nitric Oxide) can irritate the lungs, reduce lung function and increase vulnerability to allergens for those with asthma. Nitric Oxide is seen in smog and can react with other substances to harm human health.

O3 (Ozone) is the main component of smog and has been associated with eye, nose, and throat irritations. Ozone also causes shortness of breath, aggravation of respiratory conditions and allergies, chronic obstructive pulmonary disease and asthma. Increased risk of cardiovascular disease is a result of exposure to O3 as well and all of these symptoms affect the young and the elderly that live near cities and/or with acute illnesses.

Carbon monoxide can significantly impact human health by entering the blood stream through the lungs and therefore inhibiting the blood's capacity to carry oxygen to organs and tissues. CO is harmful to those with heart disease and those with respiratory conditions. It also affects healthy individuals by reducing their exercise capacity, visual perception, manual dexterity, learning functions, and ability to perform complex tasks.

SO2 (Sulfur Dioxide) can harm human respiratory systems by triggering respiratory illness and worsening existing respiratory and cardiovascular diseases. Sulfur dioxide is seen in smog and can react with other substances like nitric oxide to harm human health. When SO2 transforms into sulfate particles and combines with other particles in the atmosphere it can contribute to the formation of particulate matter (PM2.5) which harms human health as well as the environment. Sensitive populations such as children, the elderly, those with asthma are more susceptible to the affects of SO2.

NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto \[\@citestatscanNO2\]. Children are more vulnerable to these effects of respiratory damage \[\@citestatscanNO2\]. Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function \[\@citestatscanNO2\].

@@

>>>>

Exposure to NO (Nitric Oxide) can irritate the lungs, reduce lung function and increase vulnerability to allergens for those with asthma [@citeairquality]. Nitric Oxide is seen in smog and can react with other substances to harm human health.

O3 (Ozone) is the main component of smog and has been associated with eye, nose, and throat irritations [@citeairquality]. Ozone also causes shortness of breath, aggravation of respiratory conditions and allergies, chronic obstructive pulmonary disease and asthma [@citeairquality]. Increased risk of cardiovascular disease is a result of exposure to O3 as well and all of these symptoms affect the young and the elderly that live near cities and/or with acute illnesses [@citeairquality].

Carbon monoxide can significantly impact human health by entering the blood stream through the lungs and therefore inhibiting the blood's capacity to carry oxygen to organs and tissues [@citegovCanAirPoll]. CO is harmful to those with heart disease and those with respiratory conditions. It also affects healthy individuals by reducing their exercise capacity, visual perception, manual dexterity, learning functions, and ability to perform complex tasks [@citegovCanAirPoll].

SO2 (Sulfur Dioxide) can harm human respiratory systems by triggering respiratory illness and worsening existing respiratory and cardiovascular diseases [@citepembinainstitute]. Sulfur dioxide is seen in smog and can react with other substances like nitric oxide to harm human health [@citeconferenceboard]. When SO2 transforms into sulfate particles and combines with other particles in the atmosphere it can contribute to the formation of particulate matter (PM2.5) which harms human health as well as the environment [@citeconferenceboard]. Sensitive populations such as children, the elderly, those with asthma are more susceptible to the affects of SO2 [@citepembinainstitute].

NO2 (Nitrogen Dioxide) has consequences on human health including emergency hospitalizations for respiratory conditions for adults that live in lower income communities in Toronto \[\@citestatscan\]. Children are more vulnerable to these effects of respiratory damage \[\@citestatscan\]. Children that have been exposed to NO2 have an increased risk of asthma, wheezing, infections in the ear, nose and throat, influenza, serious colds, and reduced lung function \[\@citestatscan\].

@@

## References
